---
title: "Modeling Inequality: Adding Heterogeneous Agents to the OLG Model"
subtitle: "Part 6 of a series on converting 'Matlabによるマクロ経済モデル入門' to R"
author: "Your Name"
date: "2025-09-08"
categories: [R, Economics, OLG, Inequality, Shiny]
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
    self-contained: true
---

## Introduction: Beyond the "Average" Household

In our series so far, we've built a powerful model of the economy. However, it has a significant simplification: it assumes everyone within a generation is identical. The "representative agent" in our model is an average of everyone, but as the saying goes, if you have one foot in a bucket of ice and the other in a bucket of hot coals, on average you're comfortable—but the average is misleading.

The real world is characterized by **heterogeneity**. People have different skills, earn different wages, and accumulate different amounts of wealth. These differences are not just details; they are central to understanding how economic policies truly affect society. A policy that is beneficial on average might still create significant winners and losers.

In this post, we will make a major upgrade to our OLG model by introducing **heterogeneous agents**. We will move from a model with one type of household to one with multiple, distinct income groups. This will allow us to analyze the distributional consequences of economic policies and explore the dynamics of inequality.

## The "Why" and "How" of Heterogeneity

The textbook we've been following, "Matlabによるマクロ経済モデル入門," introduces this concept in Chapter 7. The core idea is to segment the population into different groups based on their lifetime earning potential.

**Why is this important?**

1.  **Analyzing Inequality:** It's impossible to talk about inequality with a model that assumes everyone is the same. By including different income groups, we can track metrics like the **Gini coefficient** or the wealth gap between the rich and the poor.
2.  **More Realistic Macroeconomics:** Different income groups behave differently. High-income households, for example, typically save a larger fraction of their income. This means that the overall national savings rate depends not just on the average income, but on the *distribution* of that income.
3.  **Distributional Effects of Policy:** A flat consumption tax is felt more heavily by low-income households who consume a larger share of their income. A change in pension benefits might be a minor issue for a wealthy retiree but a critical one for a poor retiree. Our new model can capture these effects.

**How do we implement it?**

The key is to create different **labor efficiency profiles (`SL`)** for each group. A "high-skill" group will have a higher efficiency profile, meaning they earn a higher wage at every age, while a "low-skill" group will have a lower profile. We then simulate the lifecycle decisions for *each group separately* before aggregating them to get the macroeconomic totals.

## Step 1: Modifying the R Code for Multiple Household Types

The biggest change to our R code is that we now need to loop through our different household types. Instead of calculating one path for consumption and assets, we need to calculate a set of paths for each group.

Let's say we define two groups: `group_1` (low-income) and `group_2` (high-income).

```{r}
#| label: heterogeneity-code
#| code-summary: "Conceptual R Code for Heterogeneous Agents"
#| eval: false
#| echo: true

# --- Define different labor efficiency profiles for each group ---
SL_group1 <- numeric(IDIE)
SL_group2 <- numeric(IDIE)
for (J in 1:IRET) {
  # Low-income profile (e.g., lower intercept and slope)
  SL_group1[J] <- 1.0 + 0.10 * J - 0.0020 * J^2 
  # High-income profile (e.g., higher intercept and slope)
  SL_group2[J] <- 1.8 + 0.15 * J - 0.0025 * J^2 
}

# --- The main simulation loop now needs an outer loop for groups ---

# Initialize matrices to hold results for each group
C_group1 <- matrix(0, nrow = ITER1, ncol = IDIE)
AA_group1 <- matrix(0, nrow = ITER1, ncol = IDIE)
C_group2 <- matrix(0, nrow = ITER1, ncol = IDIE)
AA_group2 <- matrix(0, nrow = ITER1, ncol = IDIE)

# Solve the household problem for each group
# (This would be inside the main `while` convergence loop)
# ...
# Solve for Group 1 using SL_group1 -> get C_group1, AA_group1
# ...
# Solve for Group 2 using SL_group2 -> get C_group2, AA_group2
# ...

# --- Macroeconomic aggregation now sums across groups ---
# Let's say Group 1 is 60% of the population and Group 2 is 40%
pop_share_g1 <- 0.6
pop_share_g2 <- 0.4

# Aggregate Consumption (AGC)
AGC[I] <- (sum(C_group1[...]) * pop_share_g1) + (sum(C_group2[...]) * pop_share_g2)

# Aggregate Assets (PASET)
PASET[I] <- (sum(AA_group1[...]) * pop_share_g1) + (sum(AA_group2[...]) * pop_share_g2)

# ... and so on for other aggregate variables.
```

## Step 2: Upgrading the Shiny App to Visualize Inequality

With our model now producing results for different groups, we need to upgrade our Shiny app to visualize them. A great way to do this is to add a new tab to our dashboard specifically for distributional analysis.

We can add two powerful new plots:

1.  **Welfare by Group:** A line chart comparing the lifetime welfare (Equivalent Variation) of the high-income and low-income groups over time. This will immediately show us who wins and who loses from a policy change.
2.  **Gini Coefficient:** A plot showing the evolution of the Gini coefficient for income or wealth. This gives us a standard, summary measure of overall inequality in the economy.

### New "Inequality Analysis" Tab in Shiny

Here's how we could structure the UI and server logic for the new tab.

```{r}
#| label: shiny-inequality-tab
#| code-summary: "New Shiny UI and Server Logic for Inequality"
#| eval: false
#| echo: true

# --- In the Quarto YAML Header ---
# format:
#   dashboard:
#     tabs:
#       - title: "Macro Results"
#         icon: "bar-chart"
#       - title: "Inequality Analysis"
#         icon: "people"

# --- In the UI (as a new tab) ---
# {r}
#| panel: tabset
#| title: "Inequality Analysis"

# ... (code to render the new plots) ...

# --- In the Server Logic ---
# renderPlot({
#   ... (run the heterogeneous simulation) ...
#
#   # Calculate welfare for each group
#   EVRT_group1 <- calculate_welfare(C_group1, ...)
#   EVRT_group2 <- calculate_welfare(C_group2, ...)
#
#   # Combine into a single data frame for plotting
#   welfare_df <- data.frame(
#     Generation = 1:ITER1,
#     Welfare = c(EVRT_group1, EVRT_group2),
#     Group = rep(c("Low-Income", "High-Income"), each = ITER1)
#   )
#
#   # Plot welfare by group
#   ggplot(welfare_df, aes(x = Generation, y = Welfare, color = Group)) +
#     geom_line() +
#     labs(title = "Lifetime Welfare by Income Group") +
#     scale_color_manual(values = c("Low-Income" = "#D55E00", "High-Income" = "#0072B2")) +
#     theme_minimal()
#
#   # Calculate and plot Gini coefficient...
# })
```

## Conclusion: A More Nuanced View

By introducing heterogeneous agents, we have transformed our model from a tool that analyzes the "average" effect into one that can speak to the critical issue of **distribution and inequality**. This is a significant step towards building a more realistic and policy-relevant simulation. It allows us to ask nuanced questions and understand that the costs and benefits of economic policies are rarely shared equally.

This upgrade showcases the power and flexibility of our R-based framework. Having built the model from the ground up, we can now extend and adapt it to explore ever more complex and interesting questions.
