---
title: "Going Global: An Open-Economy OLG Model with International Capital Flows"
subtitle: "Part 7 of a series on converting 'Matlabによるマクロ経済モデル入門' to R"
date: "2025-10-09"
categories: [R, Economics, OLG, International Finance, Shiny]
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
    self-contained: true
---

## Introduction: Beyond Domestic Borders

In our series so far, we have built a sophisticated model of a single, **closed economy**. In our simulation, all the savings generated by households must be invested within the country's own borders. While this is a useful simplification for understanding the basic mechanics of an OLG model, it misses a crucial feature of the modern world: **globalization**.

Today, capital flows freely across borders. The savings of a household in Japan can be used to finance a new factory in Vietnam. A German company can borrow from American investors to fund its research and development. These international capital flows link economies together, creating both opportunities and new challenges.

In this post, we will undertake a major upgrade to our model by opening it up to the rest of the world. We will transform our single-country model into a **two-country, open-economy OLG model**. This will allow us to analyze fascinating questions at the intersection of demography and international finance, drawing on the concepts from Chapter 9 of our textbook, "Matlabによるマクロ経済モデル入門."

## The Economics of an Open Economy

What happens when we allow capital to move freely between countries? The central mechanism is the **equalization of returns**.

-   **The Logic:** If the return on investment (the interest rate) is higher in Country A than in Country B, rational investors in Country B will move their savings to Country A to get the higher return. This outflow of capital from Country B will make capital scarcer there, driving up its interest rate. Conversely, the inflow of capital to Country A will make capital more abundant, driving down its interest rate. This process continues until the risk-adjusted return on capital is the same everywhere.
-   **The Implication:** In our model, this means we no longer have a separate interest rate for each country. Instead, there is a single **world interest rate**.

This seemingly small change has profound consequences:

1.  **Decoupling of Savings and Investment:** In a closed economy, a country's investment is strictly limited by its own domestic savings. In an open economy, a country with low savings can still have high investment by **borrowing from abroad**.
2.  **Current Account Imbalances:** This borrowing and lending between countries is tracked by the **current account**. A country that borrows from the rest of the world runs a current account deficit, while a country that lends runs a current account surplus.
3.  **New Transmission Channels:** A demographic shock in one country (like Japan's aging) can now affect other countries. For example, if Japan's savings rate falls, it might start borrowing from a younger, high-saving country, affecting the world interest rate and investment patterns globally.

## Step 1: Modifying the R Code for a Two-Country World

To build a two-country model, we essentially need to simulate two economies simultaneously and link them through a global capital market.

Let's call them "Country H" (Home) and "Country F" (Foreign).

1.  **Duplicate the Structure:** We will need to create and track all variables (consumption, assets, labor, etc.) for both countries. For example, we'll have `C_H`, `AA_H`, `C_F`, and `AA_F`.
2.  **Introduce a World Interest Rate:** Instead of calculating a separate interest rate for each country, our simulation's convergence loop will now need to solve for a single `world_R` that clears the *global* market for capital.
3.  **Track International Positions:** We need to add a new variable to track the **Net Foreign Asset (NFA)** position of each country. The NFA of Country H is the value of its assets held in Country F, minus the value of Country F's assets held in Country H. The change in NFA from one year to the next is the current account balance.

```{r}
#| label: open-economy-code
#| code-summary: "Conceptual R Code for a Two-Country Model"
#| eval: false
#| echo: true

# --- The main simulation loop now solves for a world interest rate ---

while (DIF > 0.001) {
  
  # Make a guess for the path of the world interest rate, world_R
  
  # --- Solve for Country H ---
  # Given world_R, solve the lifecycle problem for all households in Country H
  # Aggregate to find Country H's total desired savings (PA_H)
  
  # --- Solve for Country F ---
  # Given world_R, solve the lifecycle problem for all households in Country F
  # (Country F can have different demographics, e.g., a younger population)
  # Aggregate to find Country F's total desired savings (PA_F)
  
  # --- Global Market Clearing ---
  # Total world savings must equal total world capital stock
  Total_World_Savings <- PA_H + PA_F
  Total_World_Capital <- K_H + K_F
  
  # If savings > capital, the world interest rate is too high. Adjust guess for world_R down.
  # If savings < capital, the world interest rate is too low. Adjust guess for world_R up.
  
  # ... (update guess and repeat until converged) ...
}

# --- After convergence, calculate the Current Account ---
# For Country H:
# Investment_H <- K_H_next_period - (1 - RDEP) * K_H
# Savings_H <- PA_H
# Current_Account_H <- Savings_H - Investment_H
```

## Step 2: Upgrading the Shiny App for Global Analysis

Our Shiny app needs to be updated to handle this richer, two-country world. A tabbed interface becomes essential here.

1.  **Country Selection:** We can add a `selectInput` control to allow the user to choose whether they are viewing results for "Country H" or "Country F".
2.  **New International Plots:** We'll add a new tab for "International Dynamics" to display plots that only make sense in an open economy.

### New "International Dynamics" Tab in Shiny

Here's how the new UI and server logic might look.

```{r}
#| label: shiny-open-economy-tab
#| code-summary: "New Shiny UI for the Open-Economy Model"
#| eval: false
#| echo: true

# --- In the UI ---
# {r}
#| panel: sidebar

# Add a control to select the country to view
selectInput("country_view", "Select Country to Display:", 
            choices = c("Home Country", "Foreign Country"))

# ... (other sliders for policy) ...

# {r}
#| panel: tabset

# --- Tab 1: Macro Results ---
# (Plots for Savings Rate, K/L Ratio, etc., for the selected country)

# --- Tab 2: International Dynamics ---
# plotOutput("international_plots")


# --- In the Server Logic ---
# renderPlot({
#   ... (run the two-country simulation) ...
#
#   # Calculate the Current Account and Net Foreign Asset position
#   Current_Account_H <- Savings_H - Investment_H
#   NFA_H <- cumsum(Current_Account_H)
#
#   # Create a data frame for plotting
#   international_df <- data.frame(
#     Period = 1:ITER2,
#     CurrentAccount = Current_Account_H / GDP_H, # As % of GDP
#     NFA = NFA_H / GDP_H
#   )
#
#   # Plot 1: Current Account Balance
#   p1 <- ggplot(international_df, aes(x = Period, y = CurrentAccount)) +
#     geom_line() + geom_hline(yintercept = 0, linetype = "dashed") +
#     labs(title = "Current Account Balance (% of GDP)")
#
#   # Plot 2: Net Foreign Asset Position
#   p2 <- ggplot(international_df, aes(x = Period, y = NFA)) +
#     geom_line() + geom_hline(yintercept = 0, linetype = "dashed") +
#     labs(title = "Net Foreign Asset Position (% of GDP)")
#
#   grid.arrange(p1, p2, nrow = 2)
# })
```

## Conclusion: A Window on the World

By opening our model to international capital flows, we have created a powerful tool for understanding the global implications of domestic policies and demographic trends. It allows us to explore how an aging population in one country can lead to capital flowing abroad, affecting interest rates and investment opportunities for everyone.

This upgrade represents a significant leap in the complexity and realism of our simulation. It provides a framework for analyzing some of the most important macroeconomic issues of our time, from the sustainability of trade imbalances to the global effects of pension reform.
